Spin canonical divisors of log stacky curves

### 2015-06-27 stacky_write_up/spin_cring.tex
section 3 adding points theorems ammended

Furthermore, if we define $U_i$ and $V$ as in the lemma's statement and
\[
	T = (\bigcup_{i=1}^{n-1} U_i) \cup V,
\]
then we observe that $S$ and $T$ satisfy the hypothesis of Lemma \ref{lem:relations_from_generators_induction}, so we immediately prove part (b) of this lemma.

@@

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in
		\{3, 4\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements
		$y_3, y_4,$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4]$ with $grlex$ so that $y_4 \prec 
		y_3$
		and equip the ring $k[y_3, y_4, x]$ with the block 
		order so that $R = k[y_3, y_4, x]/I$. Let,
		\begin{align*}
			\initial_\prec(I) &= \initial_\prec(I')k[x, y_3, y_4] \\
			&+\langle y_4 x_j \mid 2 \leq j \leq m \rangle \\
			&+\langle y_4^2.\rangle
		\end{align*}
	%\item[(c)] Any set of minimal generators for $I'$ together with 
		%any set of relations with leading terms as in (b) minimally 
		%generate $I$.
\end{enumerate}
\end{lem}

\begin{proof}
The assumptions on $g,\sigma$ imply $H^0(\sx, 3L), H^0(\sx, 4L)$ are both base point free: If $g \geq 2$ then $\deg 3L > 2g - 1$ and $\deg 4L > 2g - 1$, so $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. If $g = 1$, we assume $\deg 3L \geq 2 > 2g - 1$, so we also have $\deg 4L \geq 2 > 2g - 1,$ so again $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. Therefore, general elements $y_3$ and $y_4$ satisfy $-\ord
_P(y_i) = 1$ by Riemann-Roch, and the assumptions on $\sigma$. \todo{Understand why/whether we need to cite Lemma 5.4.7.}

@@

\begin{align}
\label{eqn:sat_two_add_generator}
		\{ y_3^ax_2^b x_3^\epsilon \mid a \geq 0, b 
		\geq 0, \epsilon \in \{0, 1\}\} \cup \{ y_3^ay_4 \mid a \geq 0 \}
\end{align}

@@

For any $x_j$, notice that $-\ord_P(y_4x_j)\le 1$. If it is non-positive, then it lies in $R_{\halfcan'}$ immediately giving us a relation with initial term $y_4x_j$. Otherwise, 
we can choose $A_j \in k$ and $w_j\in R_{\halfcan'}$ such that $y_4x_j -A_jy_3w_j$ has no pole at $P$ and thus lies in $R_{\halfcan'}$ giving us a desired relation. 
Similarly, we can find $B, C \in k$ such that $y_4^2 + B y_3^2 x_2+ Cy_3x_2x_3$ has no pole at $P$ and is thus in $R_D$. In both cases we induce a relation, with initial term described in the lemma's statement. Noting that any monomial is either a multiple of one of these initial terms, is a basis element (described above), or is in $R_D$, we see that these monomials in fact generate the initial ideal of $I$ over $\initial(I')$.

%Finally, (c) follows immediately from Lemma ~\ref{lem:minimal_quadratic}.
\end{proof}

\begin{lem}
\label{lem:sat-3}
\todo{Make explicit assumptions on genus.}
Suppose $L'$ is a log half-canonical divisor of $\sx'$ with course
space $X$ of genus 0 such that $\sat(\Eff(\halfcan')) = 3$ with $R_
{\halfcan'} = k[x_3, x_4 , x_5, \ldots, x_m]/I'$. Let $L = L' + \frac
{1}{3}P$. Suppose $\deg x_i = i$ for $i \in \{3, 4, 5\}$ and that
the ordering on $k[x]$ satisfies
\begin{align*}
	\ord_{x_3}(f) < \ord_{x_3}(h) \implies f \prec h.
\end{align*}

@@

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx, iL)$ for $i \in \{3,
		4,5\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3, y_4,$ and $y_5$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4, y_5]$ with $grlex$ so that $y_5 \prec 
		y_4 \prec y_3$
		and equip the ring $k[y_3, y_4, y_5, x]$ with the block 
		order so that $R = k[y_3, y_4, y_5, x]/I$. Let,
		\begin{align*}
			\initial_\prec(I) &= \initial_\prec(I) k[x, y_3, y_4, y_5] \\
			&+ \langle y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\rangle \\
			&+ \langle y_i y_k \mid 4 \leq i \leq j \leq 5.\rangle 
		\end{align*}
	\item[(c)] Any set of generators for $I'$ together with 
		any set of relations with leading terms as in (b) 
		generate $I$.
\end{enumerate}
\end{lem}

\begin{proof}
First, note that since we are in genus 0, we can choose the
elements $y_3, y_4,$ and $y_5$ are general and have $- \ord_P(y_i) =
1$ by Riemann Roch. \todo{Understand why/whether we need to cite 
Lemma 5.4.7.} To complete the proof of (a) we only need show these
elements generate $R$ over $R'$. We can first see that the
following is a $k$ basis for $R$ over $R'$:

\begin{align}
\label{eqn:add_one_generator}
	\begin{split}
		&\langle y_3^ax_3^b x_4^\epsilon x_5^{\epsilon'} \mid a \geq 0, b 
		\geq 0,(\epsilon,\epsilon') \in \{(0,0),(0,1),(1,0)\} \rangle \\
		&+\langle y_3^ay_4, y_3^by_5 \mid a \geq 0, b \geq 0 \rangle
	\end{split}
\end{align}

\noindent
To see these generate $R'$ over $R$, note that $\dim_k R'_d - \dim_k R
_d = \lfloor \frac{d}{3} \rfloor,$ by Riemann--Roch. So, it
suffices to show that we have precisely $\lfloor \frac{d}{3} \rfloor
,$ elements of degree $d$ in the claimed basis of
~\ref{eqn:add_one_generator}. Indeed, letting $a = \lfloor \frac{d}{3}
\rfloor $ and $b = d \bmod 3$, we have that the elements

@@

\noindent
Of course, the initial terms of these elements are precisely the
generators of $J$, completing (b).

>>>>

Furthermore, if we define $U_i$ and $V$ as in the lemma's statement and
\[
	T = (\bigcup_{i=1}^{n-1} U_i) \cup V,
\]
then we observe that $S$ and $T$ satisfy the hypothesis of Lemma \ref{lem:relations_from_generators_induction}, so parts (b) and (c) of the lemma follow immediately

@@

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in
		\{3, 4\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements
		$y_3, y_4,$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4]$ with $grlex$ so that $y_4 \prec 
		y_3$
		and equip the ring $k[y_3, y_4, x]$ with the block 
		order so that $R = k[y_3, y_4, x]/I$. Let,
		\begin{align*}
			\initial_\prec(I) &= \initial_\prec(I')k[x, y_3, y_4] \\
			&+\langle y_4 x_j \mid 2 \leq j \leq m \rangle \\
			&+\langle y_4^2 \rangle.
		\end{align*}
	%\item[(c)] Any set of minimal generators for $I'$ together with 
		%any set of relations with leading terms as in (b) minimally 
		%generate $I$.
\end{enumerate}
\end{lem}

\begin{proof}
The assumptions on $g$ imply $H^0(\sx, 3L), H^0(\sx, 4L)$ are both base point free: If $g \geq 2$ then $\deg 3L > 2g - 1$ and $\deg 4L > 2g - 1$, so $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. If $g = 1$, we assume $\deg 3L \geq 2 > 2g - 1$, so we also have $\deg 4L \geq 2 > 2g - 1,$ so again $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. Therefore, general elements $y_3$ and $y_4$ satisfy $-\ord
_P(y_i) = 1$ by Riemann-Roch. \todo{Understand why/whether we need to cite Lemma 5.4.7.}

@@

\begin{align}
\label{eqn:sat_two_add_generator}
	S =	& \; \{ y_3^ax_2^b x_3^\epsilon \mid a \geq 0, b 
		\geq 0, \epsilon \in \{0, 1\}\} \\
	\cup	& \; \{ y_3^ay_4 \mid a \geq 0 \}
\end{align}

@@

Letting
\begin{align*}
	T =   &\; \{ y_4 x_j \mid 2 \leq j \leq m \} \\
	\cup &\; \{ y_4^2 \}
\end{align*}
part (b) follows immediately from lemma \ref{lem:relations_from_generators_induction}

\iffalse{
For any $x_j$, notice that $-\ord_P(y_4x_j)\le 1$. If it is non-positive, then it lies in $R_{\halfcan'}$ immediately giving us a relation with initial term $y_4x_j$. Otherwise, 
we can choose $A_j \in k$ and $w_j\in R_{\halfcan'}$ such that $y_4x_j -A_jy_3w_j$ has no pole at $P$ and thus lies in $R_{\halfcan'}$ giving us a desired relation. 
Similarly, we can find $B, C \in k$ such that $y_4^2 + B y_3^2 x_2+ Cy_3x_2x_3$ has no pole at $P$ and is thus in $R_D$. In both cases we induce a relation, with initial term described in the lemma's statement. Noting that any monomial is either a multiple of one of these initial terms, is a basis element (described above), or is in $R_D$, we see that these monomials in fact generate the initial ideal of $I$ over $\initial(I')$.
%Finally, (c) follows immediately from Lemma ~\ref{lem:minimal_quadratic}.
}\fi
\end{proof}

\begin{lem}
\label{lem:sat-3}
\todo{Make explicit assumptions on genus.}
Suppose $L'$ is a log half-canonical divisor of $\sx'$ with course
space $X$ of genus 0 such that $\sat(\Eff(\halfcan')) = 3$ and $R_
{\halfcan'} = k[x_3, x_4 , x_5, \ldots, x_m]/I'$. Let $L = L' + \frac
{1}{3}P$. Suppose $\deg x_i = i$ for $i \in \{3, 4, 5\}$ and that
the ordering on $k[x]$ satisfies
\begin{align*}
	\ord_{x_3}(f) < \ord_{x_3}(h) \implies f \prec h.
\end{align*}

@@

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx, iL)$ for $i \in \{3,
		4,5\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3, y_4,$ and $y_5$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4, y_5]$ with $grlex$ so that $y_5 \prec 
		y_4 \prec y_3$
		and equip the ring $k[y_3, y_4, y_5, x]$ with the block 
		order so that $R = k[y_3, y_4, y_5, x]/I$. Let,
		\begin{align*}
			\initial_\prec(I) &= \initial_\prec(I) k[x, y_3, y_4, y_5] \\
			&+ \langle y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\rangle \\
			&+ \langle y_i y_k \mid 4 \leq i \leq j \leq 5\rangle.
		\end{align*}
	%\item[(c)] Any set of generators for $I'$ together with 
	%	any set of relations with leading terms as in (b) 
	%	generate $I$.
\end{enumerate}
\end{lem}

\begin{proof}
Since we are genus 0, general
elements $y_3, y_4,$ and $y_5$ in weights 3, 4, and 5 respectively satisfy $- \ord_P(y_i) =
1$. \todo{Understand why/whether we need to cite 
Lemma 5.4.7.}  We see by pole order considerations that

\begin{align}
\label{eqn:add_one_generator}
	\begin{split}
	S=	&\{ y_3^ax_3^b x_4^\epsilon x_5^{\epsilon'} \mid a \geq 0, b 
		\geq 0,(\epsilon,\epsilon') \in \{(0,0),(0,1),(1,0)\} \} \\
	\cup \;&\{y_3^ay_4, y_3^by_5 \mid a \geq 0, b \geq 0 \}
	\end{split}
\end{align}

\noindent forms a $k$ basis for $R$ over $R'$, which concludes part (a) of the proof.

Letting
\begin{align*}
	T = & \{ y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\} \\
	   + & \{ y_i y_k \mid 4 \leq i \leq j \leq 5\} 
\end{align*}
we see that $S$ and $T$ along with $\prec$ satisfy the hypothesis of Lemma \ref{lem:relations_from_generators_induction}, immediately implying part (b) of this proof.

\iffalse{
To see these generate $R'$ over $R$, note that $\dim_k R'_d - \dim_k R
_d = \lfloor \frac{d}{3} \rfloor,$ by Riemann--Roch. So, it
suffices to show that we have precisely $\lfloor \frac{d}{3} \rfloor
,$ elements of degree $d$ in the claimed basis of
~\ref{eqn:add_one_generator}. Indeed, letting $a = \lfloor \frac{d}{3}
\rfloor $ and $b = d \bmod 3$, we have that the elements

@@

\noindent
Of course, the initial terms of these elements are precisely the
generators of $J$, completing (b).
}\fi
